Repairing cooperation through a cognitive intervention in the repeated Trust Game

### 2023-06-30 article.Rmd
Final edits

---
title: "Repairing cooperation through a cognitive intervention in the repeated Trust Game"
author:
  - Ismail Guennouni
  - Maarten Speekenbrink
  - Quentin Huys
  - Samuel Dupret

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011].

@@

@@

cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

<!-- MS: double check statistics; I get different df when using KR in mixed -->
Mixed-effects analysis shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f` <!--$F(1,315.31) = 9.52$, $p = .002$-->, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game-number (pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f` <!--$F(1,314.2) = 26.9$, $p < .001$-->. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:violinPanels}.A). This indicates the intervention was effective in increasing cooperative behaviour. There was also a significant main effect of Investment, `r papaja::apa_print(mod_returns_pct)$full_result$inv_scaled` <!--$F(1,9208) = 373.6$, $p < 0.001$-->, such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, `r papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled` <!--$F(1,9207) = 45.38$, $p < .001$-->, indicates the positive effect of investment on percentage returns was greater in the control than intervention condition. <!-- This suggests that the higher returns we see in the intervention group are not simply due to higher investments.--> There was also an Investment by Game-number interaction, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled`. Finally, we find a three way interaction between Game-number, Condition and Investment, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled`, <!--($F(1,8988)= 24.6, p < 0.001$),--> showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game-number.

@@

We next analysed HMM agent investments via a linear-mixed effects model with fixed effects for Condition (intervention or control), Game-number (pre or post intervention), as well as interaction between Condition and Game-number, and participant-wise random intercepts and random slopes for Game-number. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game-number, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:violinPanels}.B, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

Taken together, we find that participants in the control condition sent *lower* percentage returns in the second game, despite the HMM investor sending on average higher investments in the second game. Those in the intervention group returned *higher* percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention group compared to the control group were not purely driven by reciprocity towards higher investments, since we find a Condition by Game-number interaction whilst controlling for investment in the model, and a reduced effect of Investment in the Intervention condition. <!-- since we find this interaction effect whilst controlling for investments in our model of participants returns.-->

@@

We next analysed returns separately for rounds prior to the pre-programmed defection (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed effects model as before to returns before defection, largely replicates results over all trials. Full results are provided in the supplement. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-defection trials between the two phases of the RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`.

@@

The intervention focused on repairing cooperation after a "transgressive action" by the investor. To assess whether the intervention succeeded, we next considered the returns by participants after the pre-programmed transgression (Figure \ref{fig:violinPanels}.D). <!-- We fit the same linear mixed effects model to percentage returns as the one for the full data while restricting the data to rounds 12 to 15 in the first game, and 13 to 15 in the second game. These trials represent all trials after the pre-programmed defection of the investor (low investment of $2$) until the end of the game (round 15).
-->
<!-- Figure \ref{fig:violinPanels}.D shows the marginal means of trustee returns and their distribution by Game-number and Condition.--> A mixed-model analysis shows similar results to those of the model applied to all returns (see SI for full results). The was again a significant interaction between Condition and Game-number and Condition. Participants in the control condition decreased their post-transgression returns from the first to the second phase of the RTG, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`. So, whilst the intervention did not appear to move participants to increase their returns after a transgression by the investor, it may have countered the decrease in returns showed by participants in the control condition.

@@

Participants in the intervention condition rated their emotion on valence (negative to positive) and arousal (low to high) after each investment. To assess the impact of the intervention on these emotional reactions, we used linear mixed-effects models (one for valence, and one for arousal) with fixed effects for Game-number (pre or post intervention) and Investment, as well as interaction between Investment and Game-number, with participant-wide random intercepts and random slopes for Game-number. This showed that higher investments were associated with more positive emotions, `r papaja::apa_print(mod_emo_x)$full_result$scaleinvestment`, and higher arousal, `r papaja::apa_print(mod_emo_y)$full_result$scaleinvestment`. <!-- ($F(1,4664) = 2902, p < 0.001$), and higher arousal ($F(1,4668.4) = 1919, p < 0.001$)--> In addition, the positiveness of emotion declined between the two games, `r papaja::apa_print(mod_emo_x)$full_result$gameNum_f`, as did arousal, `r papaja::apa_print(mod_emo_y)$full_result$gameNum_f`. There was no indication that the effect of the investment on either aspect of emotion was affected by the intervention, as there was no interaction between Investment and Game-number on valence, `r papaja::apa_print(mod_emo_x)$full_result$gameNum_f_scaleinvestment`, or arousal, `papaja::apa_print(mod_emo_y)$full_result$gameNum_f_scaleinvestment`. This indicates that participants in the intervention condition returned higher amounts post-intervention despite their emotional reaction to investments remaining largely the same.

@@

Participants rated the HMM investor in the second game as less cooperative (`r apa_print(pairs(emmeans::emmeans(coop_mod, c("gameNum.f"))))$full_result`), less trustworthy (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("gameNum.f"))))$full_result`), less friendly (`r apa_print(pairs(emmeans::emmeans(friend_mod, c("gameNum.f"))))$full_result`) and more selfish (`r apa_print(pairs(emmeans::emmeans(self_mod, c("gameNum.f"))))$full_result`), than the HMM investor in the first game. Participants in the intervention condition rated players higher than those in the control condition on cooperativeness (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("condition.f"))))$full_result`) and lower on selfishness (`r apa_print(pairs(emmeans::emmeans(self_mod, c("condition.f"))))$full_result`). There was no evidence for an interaction effect between Game-number and Condition on any of the attributes.

@@

We next asked whether the intervention had any discernible effect on participants' behaviour in a different, repeated Prisoner's Dilemma game. Predicting the probability of a cooperative action with a logistic mixed-effects regression model, with Condition and Phase (before or after defection trial) as fixed effects and a random intercept for participants, showed a decline in cooperation after defection by the other player, `r papaja::apa_print(ipd_mod)$full_result$phase`, but no evidence for a different cooperation rate in the intervention condition compared to the control condition, `r `papaja::apa_print(ipd_mod)$full_result$condition_f`, or a different response to defection between the conditions, `r papaja::apa_print(ipd_mod)$full_result$condition_f_phase`. <!-- We also found no difference in cooperation rates post defection trial between conditions.-->

@@

When asked whether they thought their opponent was Human or not, $40$% of participants thought they were either facing a human or were not sure of the nature of the opponent. Many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial opponent's behavior.

@@

We analyzed participants' behavior differences in the intervention versus control conditions using hidden Markov models (HMM). Five models were used: "HMM-inv" assumed transition states depended solely on investment, ignoring game number and condition. "HMM-prepost" contrasted Pre and Post Intervention. "HMM-coax" contrasted post-intervention with pre-intervention and both control conditions. "HMM-ctrl" contrasted post-control with pre-control and both intervention conditions. "HMM-full" grouped pre-control and pre-intervention as one, with separate groups for post-control and post-intervention. Models were fitted using 2 to 7 states, selecting the lowest BIC. Generally, 5-7 state models best explained the data. A likelihood ratio test compared the models' goodness of fit, contrasting the complex HMM-full model with nested models equating behavior in certain stages and conditions.

@@

In order to compare the goodness of fit of the various models, we test the relative likelihood of models using a likelihood ratio test. This procedure is useful to compare the most complex model (HMM-full, which allows for differences between pre-intervention and the two conditions post-intervention) to nested models which equate behaviour in some if the stages and conditions.

@@

Using likelihood ratio tests, we find that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCtrl[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCoax[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedPrePost[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$). This is consistent with a differentiated behavior of the trustees between all three groups: the post-intervention group, the post-control group and the pre-manipulation group.

@@

Using the HMM-full model, we can retrieve participants' return distributions based on their latent states (Figures \ref{fig:postStatesBestMod}.A) and transition probabilities between these states (Figure \ref{fig:plotTransitionsDiscGaus}). The states are ranked by mean return, with State 1 having the lowest mean return and State 5 the highest. A higher state number indicates a more pro-social policy. We focused on states related to cooperation's breakdown and repair. We compared the transition probabilities between states when the investment is low for post-control and post-intervention groups. Figure \ref{fig:plotTransitionsDiscGaus} suggests that the intervention group is more forgiving of low investments, as they are less likely to shift to an anti-social state when faced with defection compared to the control group.

@@

To quantitatively explore the differences in transition probabilities between the control and intervention conditions, we can estimate from the model, using local decoding methods from the depmixS4 package [@visser_depmixs4_2021], the most likely posterior state of the trustee participants by round given the actions they have taken. Figure \ref{fig:postStatesBestMod}.B shows that participants were more likely to be in a lower return state in the control condition compared to the intervention condition both pre and post defection. For instance, in round 5, state 1 was the most likely posterior state for only $7$% of participants in the intervention condition compared to $24$% in the control condition ($\chi^2(1) = 8.26, p < 0.01$). For the post-defection trial after the intervention (round 14), state 1 was the most likely state for only $22$% of participants in the intervention condition compared to $43$% in the control condition ($\chi^2(1) = 14.70, p < 0.001$).

The posteriors also suggest that a non-negligible proportion of participants in the intervention condition did not exhibit a behaviour consistent with the goal of the intervention as they were still best fit by low-return states post intervention. For instance, focusing on round 13 post defection $30.2$% of those in the intervention condition were most likely to be in the least pro-social state 1. These differences can be seen as an indication of important heterogeneity in the effectiveness of the intervention.

@@

In this experiment, we made human participants face artificial computer agents endowed with the ability to transition between latent states and react to the participants' returns. The number of states, the policy in each state as well as the way these agents transitioned between states was based on estimating a hidden Markov model to behaviour from real human participants. On average, we saw the emergence of cooperative behavior with investment and returns in line with what is reported in human dyadic interaction in the repeated trust game [@charness_investment_2008; @fiedler_social_2011]. The emergence of cooperative behavior and participant's uncertainty about whether they were facing human or artificial opponents, point to the potential of these agents to mimic human behavior in economic games whilst offering a higher degree of experimental control.

The intervention's aim was to articulate the effect of acting on impulse in case of a transgressive action from the investor in the form of a one-off low investment. When the agent defected as programmed, yielding a situation similar to the one presented in the intervention, the intervention led to higher percentage returns post defection compared to the control group, as intended. Participants emotional reactions to the received investment, we similar pre and post-intervention, so the higher returns were produced despite participants having the same emotional reaction. This might indicate that the intervention achieved its goal of encouraging participants to respond in a non-impulsive, considered way, overriding the emotional urge to retaliate.

What was also notable was that the intervention led to generally higher returns post-intervention, even before the defection trial. This could be for a number of reasons. One possibility is that participants simply learned that by returning more, the investment on the next trial would be higher. However, this explanation is unlikely, as we did not see a similar increase in percentage returns in the control condition. Another possibility is that, on average, the second player they faced invested more, which prompted positive reciprocity in the human trustees. As all players were programmed in the same way, any difference would be due to participant's actions. Further, participants rated both players similarly on relevant attributes of cooperativeness and trust. The higher returns are thus unlikely to be driven by different beliefs about the investor. A likely explanation for overall higher trustworthiness, as measured by higher percentage return post investment, is that participants, on average, interpreted the intervention message as an argument for more pro-social behavior, irrespective of the investor's actions.

It is also noteworthy that there were important individual differences in the percentage return changes post vs. pre intervention, which can be seen as a proxy for the intervention effectiveness. Some participants might not have been convinced by the intervention's message and decided to reduce their returns both pre and post defection in the second trust game, while others increased their returns in both phases. This raises important questions for the measurement of intervention effectiveness. Recent work has shed light on the important heterogeneity inherent in how disorders are categorised: This heterogeneity arises from the view that mental health problems should be viewed as complex systems, or interactions between neuro-computational processes and socio-environmental contexts evolving over time [@fried_moving_2017]. This view was used to justify computational psychiatry's difficulty in establishing differential and reliable predictors of likely treatment responses [@hitchcock_computational_2022]. But if a healthy group's reaction to a relatively explicit intervention is itself heterogeneous as we have shown in this experiment, then the issue of variable treatment responses might the result of the interaction of two sources of variability: the phenotyping of the disorder as well as the phenomenological aspects of the intervention itself. As such, a rigorous exploration of the determinants of inter-individual differences to an intervention in the general patient population is required.

>>>>

---
title: "Repairing cooperation through a cognitive intervention in the repeated Trust Game"
author:
  - Ismail Guennouni
  - Quentin Huys
  - Samuel Dupret
  - Maarten Speekenbrink

@@

@@

```
```{r, include=FALSE, cache=TRUE}

@@

cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```
```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

<!-- MS: double check statistics; I get different df when using KR in mixed -->

Average investments and returns prior to the "defection round" (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`<!--$F(1,315.31) = 9.52$, $p = .002$-->, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`<!--$F(1,314.2) = 26.9$, $p < .001$-->. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:violinPanels}.A). This indicates the intervention was effective in increasing cooperative behaviour. There was also a significant main effect of Investment, `r papaja::apa_print(mod_returns_pct)$full_result$inv_scaled`<!--$F(1,9208) = 373.6$, $p < 0.001$-->, such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, `r papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled`<!--$F(1,9207) = 45.38$, $p < .001$-->, indicates the positive effect of investment on percentage returns was greater in the control than intervention condition. <!-- This suggests that the higher returns we see in the intervention group are not simply due to higher investments.--> There was also an Investment by Game interaction, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled`. Finally, we find a three way interaction between Game, Condition and Investment, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled`<!--($F(1,8988)= 24.6, p < 0.001$),-->, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

@@

We next analysed HMM agent investments with a linear-mixed effects model with fixed effects for Condition (intervention vs control), Game (pre vs post intervention), as well as interaction between Condition and Game, and participant-wise random intercepts and random slopes for Game. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:violinPanels}.B, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

<!-- Taken together, we find that participants in the control condition sent lower percentage returns in the second game, despite the HMM investor sending on average higher investments in the second game. Those in the intervention group returned higher percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention compared to the control condition were not purely driven by reciprocity towards higher investments, since we find a Condition by Game interaction whilst controlling for investment in the model, and a reduced effect of investment in the intervention condition. <!-- since we find this interaction effect whilst controlling for investments in our model of participants returns.-->

@@

We next analysed returns separately for rounds prior to the pre-programmed transgression (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed effects model as before to returns before the transgression largely replicates the results over all trials. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-transgression trials between the first and second RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`. Full results are provided in the SI.

@@

<!-- The intervention focused on repairing cooperation after a "transgressive action" by the investor. To assess whether the intervention succeeded, we next considered the returns by participants after the pre-programmed transgression (Figure \ref{fig:violinPanels}.D).--> <!-- We fit the same linear mixed effects model to percentage returns as the one for the full data while restricting the data to rounds 12 to 15 in the first game, and 13 to 15 in the second game. These trials represent all trials after the pre-programmed defection of the investor (low investment of $2$) until the end of the game (round 15).
-->
<!-- Figure \ref{fig:violinPanels}.D shows the marginal means of trustee returns and their distribution by Game-number and Condition.--> Applying the same model to the returns after the transgression (see SI for full results), we again find <!-- A mixed-model analysis shows similar results to those of the model applied to all returns (see SI for full results). There was again --> a significant interaction between Condition and Game. Participants in the control condition decreased their post-transgression returns from the first to the second RTG, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no significant change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`. <!-- So, whilst the intervention did not appear to move participants to increase their returns after a transgression by the investor, it may have countered the decrease in returns showed by participants in the control condition. -->

Taken together, we find that participants in the control condition sent lower percentage returns in the second game, despite the HMM investor sending on average higher investments in the second game. Those in the intervention group returned higher percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention compared to the control condition were not purely driven by reciprocity towards higher investments, since we found a Condition by Game interaction whilst controlling for investment in the model, and a reduced effect of investment in the intervention condition. Whilst the intervention increased participants returns overall, it did not appear to increase participants' returns after the transgression by the other player. However, it may have countered the tendency to lower returns after the transgression in the second game compared to the first game, which was shown by participants in the control condition.

@@

Participants in the intervention condition rated their emotion on valence (negative to positive) and arousal (low to high) after each investment. To assess the impact of the intervention on these emotional reactions, we used linear mixed-effects models (one for valence, and one for arousal) with fixed effects for Game (pre vs post intervention) and Investment, as well as interaction between Investment and Game, with participant-wide random intercepts and random slopes for Game. This showed that higher investments were associated with more positive emotions, `r papaja::apa_print(mod_emo_x)$full_result$scaleinvestment`, and higher arousal, `r papaja::apa_print(mod_emo_y)$full_result$scaleinvestment`. <!-- ($F(1,4664) = 2902, p < 0.001$), and higher arousal ($F(1,4668.4) = 1919, p < 0.001$)--> In addition, the positiveness of emotion declined between the two games, `r papaja::apa_print(mod_emo_x)$full_result$gameNum_f`, as did arousal, `r papaja::apa_print(mod_emo_y)$full_result$gameNum_f`. There was no indication that the effect of the investment on either aspect of emotion was affected by the intervention, as there was no interaction between Investment and Game on valence, `r papaja::apa_print(mod_emo_x)$full_result$gameNum_f_scaleinvestment`, or arousal, `r papaja::apa_print(mod_emo_y)$full_result$gameNum_f_scaleinvestment`. This indicates that participants in the intervention condition returned higher amounts post-intervention, despite their emotional reaction to investments remaining largely the same.

@@

Participants rated the HMM investor in the second game as less cooperative (`r apa_print(pairs(emmeans::emmeans(coop_mod, c("gameNum.f"))))$full_result`), less trustworthy (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("gameNum.f"))))$full_result`), less friendly (`r apa_print(pairs(emmeans::emmeans(friend_mod, c("gameNum.f"))))$full_result`) and more selfish (`r apa_print(pairs(emmeans::emmeans(self_mod, c("gameNum.f"))))$full_result`), than the HMM investor in the first game. Participants in the intervention condition rated players higher than those in the control condition on cooperativeness (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("condition.f"))))$full_result`) and lower on selfishness (`r apa_print(pairs(emmeans::emmeans(self_mod, c("condition.f"))))$full_result`). There was no evidence for an interaction effect between Condition and Game on any of the attributes.

@@

We next asked whether the intervention had any discernible effect on participants' behaviour in a different, repeated Prisoner's Dilemma game. Predicting the probability of a cooperative action with a logistic mixed-effects regression model, with Condition and Phase (before or after defection trial) as fixed effects and a random intercept for participants, showed a decline in cooperation after defection by the other player, `r papaja::apa_print(ipd_mod)$full_result$phase`, but no evidence for a different cooperation rate in the intervention condition compared to the control condition, `r `r papaja::apa_print(ipd_mod)$full_result$condition_f`, or a different response to defection between the conditions, `r papaja::apa_print(ipd_mod)$full_result$condition_f_phase`. As such, there is no evidence that the intervention affected behaviour in this game. <!-- We also found no difference in cooperation rates post defection trial between conditions.-->

@@

When asked whether they thought their opponent was Human or not, $40$% of participants thought they were either facing a human or were not sure of the nature of the opponent. Many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial opponent's behaviour.

@@

We used hidden Markov models (HMMs) to further assess differences between the intervention and control condition in participants' reactions to the investor. As in the models for the investor, these HMMs assume behaviour is governed by latent states, with participants' switches between states now dependent on the investments made. We also allowed for differences between games and conditions in how investments govern state transitions. More in particular, we fitted five main models which all regressed state transition probabilities onto investments, as well as on additional contrast-coded predictors for condition and/or game number. In the most complex model (HMM-full), the transition probabilities were allowed to differ between all four combinations of game number and condition. The HMM-coax model allowed differences between post-intervention and the other three conditions (pre-intervention, pre-control, post-control) assuming these latter were the same. Similarly, the HMM-ctrl model allowed differences between post-control and the other three conditions. The HMM-prepost model allowed differences between the first and second RGT. Finally, the HMM-inv model did not allow transition probabilities to differ between conditions or game numbers, modelling them only as a function of investment. As the number of hidden states was unknown, for each model we estimated versions with 2 to 7 states, and used the BIC to select the best fitting model. Generally, best fitting models had between 5 to 7 hidden states.

@@

llratio(newFull[[6]],fittedCtrl[[6]])
llratio(newFull[[6]],fittedCoax[[6]])
llratio(newFull[[6]],fittedPrePost[[6]])
```

<!-- MS: There seems to be an issue with the likelihoods of the models. For example, for 6 states models, the likelihood of the Ctrl and Coax model are higher than for the Full model, but this is theoretically not possible as additional parameters can only increase the likelihood! -->

Focussing on models with 5 latent states (which is the optimal number for the HMM-full model), likelihood ratio tests showed that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCtrl[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCoax[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedPrePost[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$). This is consistent with state transitions differing both due to game number and condition.

@@

<!-- Using likelihood ratio tests, we find that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCtrl[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedCoax[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(40) = `r round(-2*as.numeric(logLik(fittedPrePost[[5]])) - (-2*as.numeric(logLik(newFull[[5]]))) , 2)`$, $p < .001$). This is consistent with a differentiated behavior of the trustees between all three groups: the post-intervention group, the post-control group and the pre-manipulation group. -->

@@

<!-- MS: I think Figure 6 can be moved to the supplement and perhaps Figure 6.A too. -->

<!-- Using the HMM-full model, we can retrieve participants' return distributions based on their latent states (Figures \ref{fig:postStatesBestMod}.A) and transition probabilities between these states (Figure \ref{fig:plotTransitionsDiscGaus}). The states are ranked by mean return, with State 1 having the lowest mean return and State 5 the highest. A higher state number indicates a more pro-social policy. We focused on states related to cooperation's breakdown and repair. We compared the transition probabilities between states when the investment is low for post-control and post-intervention groups. Figure \ref{fig:plotTransitionsDiscGaus} suggests that the intervention group is more forgiving of low investments, as they are less likely to shift to an anti-social state when faced with defection compared to the control group. -->

@@

Focusing on the best-fitting 5-state HMM-full model, we used a local decoding procedure to assign observations (participants' returns on trials) to latent states. The states are ordered by expected return, with state 1 having the lowest mean return and state 5 the highest. Figure \ref{fig:postStatesBestMod}.B shows that participants were more likely to be in a lower return state in the control condition compared to the intervention condition both pre and post defection. For instance, in round 5, state 1 was the most likely posterior state for only $7$% of participants in the intervention condition compared to $24$% in the control condition ($\chi^2(1) = 8.26, p < 0.01$). For the post-defection trial after the intervention (round 14), state 1 was the most likely state for only $22$% of participants in the intervention condition compared to $43$% in the control condition ($\chi^2(1) = 14.70, p < 0.001$). Whilst the posterior states indicate that the intervention was effective, a non-negligible proportion of participants in the intervention condition did not exhibit the coaxing behaviour promoted by the intervention. Directly following the low investment in round 13, $30.2$% of participants in the intervention condition were assigned to state 1 with the lowest average returns. There is thus clear heterogeneity in the effectiveness of the intervention.

<!-- To quantitatively explore the differences in transition probabilities between the control and intervention conditions, we can estimate from the model, using local decoding methods from the depmixS4 package [@visser_depmixs4_2021], the most likely posterior state of the trustee participants by round given the actions they have taken. Figure \ref{fig:postStatesBestMod}.B shows that participants were more likely to be in a lower return state in the control condition compared to the intervention condition both pre and post defection. For instance, in round 5, state 1 was the most likely posterior state for only $7$% of participants in the intervention condition compared to $24$% in the control condition ($\chi^2(1) = 8.26, p < 0.01$). For the post-defection trial after the intervention (round 14), state 1 was the most likely state for only $22$% of participants in the intervention condition compared to $43$% in the control condition ($\chi^2(1) = 14.70, p < 0.001$).

The posteriors also suggest that a non-negligible proportion of participants in the intervention condition did not exhibit a behaviour consistent with the goal of the intervention as they were still best fit by low-return states post intervention. For instance, focusing on round 13 post defection $30.2$% of those in the intervention condition were most likely to be in the least pro-social state 1. These differences can be seen as an indication of important heterogeneity in the effectiveness of the intervention.
-->

@@

In this experiment, human participants took the role of trustee in a repeated trust game (RTG) where they faced artificial computer agents who's behaviour was partly determined by participants' returns. The behaviour of the artificial investors was determined by a 3-state hidden Markov model (HMM), which was estimated from the behaviour of humans in the RTG in prior research. Overall, investments and returns by the artificial and human agents replicated that of human dyads in prior research using the RTG [@charness_investment_2008; @fiedler_social_2011]. This, together with participants' reported uncertainty about whether they faced a human or artificial investor, shows the potential of using HMM-based artificial agents to mimic human behaviour in economic games, whilst offering a higher degree of experimental control.

The aim of the cognitive intervention was to articulate the potential unwanted effects of acting on impulse after a transgressive action from the investor in the form of a one-off low investment. After the intervention, participants send back higher returns compared to before the intervention, and did not decrease their returns after a transgressive action like participants in the control condition did. The overall higher returns after the intervention occurred despite participants' emotional reactions to the investments remaining largely the same as before the intervention. This indicates the intervention achieved its goal of encouraging participants to respond in a non-impulsive and considered manner, overriding an emotional urge to retaliate.

That participants generally send higher returns to the investor after the intervention is unlikely due to a  general learning effect unrelated to the intervention, as participants in the control condition did not increase their returns. Also, as participants in the intervention and control condition faced the same HMM investor, the higher post-intervention returns are not solely due to a difference in investor behaviour. As the investor reacts to participants' returns, those who return more will generally see higher investments. But this is driven by the magnitude of their returns, not by a change in the strategy of the investor. Finally, as participants rated the first and second HMM agent similarly on attributes such as cooperativeness and trust, the increased returns are unlikely to be the result of a more favourable evaluation of the investor. Rather, we find it most likely that the increased returns are due to participants inferring from the intervention that pro-social and trustworthy behaviour may generally motivate the investor to send high investments, which can provide more beneficial outcomes to them in the long run.

The effect of the intervention was not transferred to the Repeated Prisoner's Dilemma (RPD) game. There was no difference between the intervention and control condition in the rate of cooperation, whether before or after a preprogrammed defection by the artificial agent. As the prisoner's dilemma is a popular economic game, it is possible that participants had a strong prior commitment towards the strategy they would adopt, which was not overridden by the intervention. The RPD also involves much coarser actions (cooperate vs defect) than the finer-grained returns in the RTG. This would make it more difficult to observe more subtle effects of the intervention. As such, the RPD might not be the best choice for any knowledge transfer. In any case, we can not rule out that the effects of the brief cognitive intervention, which explicitly focused on the RTG, are confined to the RTG. Future research will need to assess whether transfer to other games is possible. <!-- For those that took on the intervention message and showed coaxing behavior in the second trust game, the fact that the investor still defected in the final rounds might have reinforced the idea that not reciprocating negative behavior is a losing strategy after all.-->

There were clear individual differences in the change of returns between the pre- and post-intervention RTG, which can be seen as a proxy for the effectiveness of the intervention. Some participants may not have been convinced that coaxing via high returns was a good way to establish cooperation and decided to reduce their returns in the second trust game. Their impulse to “punish” the other player for a transgression may have been too strong to be overridden by the intervention. This was also evident from the participants' replies to a question about whether they would change their behaviour, just after receiving the intervention. Other participants responded to the intervention and increased their returns. <!-- This raises important questions for the measurement of intervention effectiveness.--> Heterogeneity in response to treatment is common in psychiatry and related fields. Such heterogeneity may reflect the complex nature of mental health problems, which may be best viewed as complex systems involving interactions between neuro-computational processes and socio-environmental contexts evolving over time [@fried_moving_2017]. This view was used to justify computational psychiatry's difficulty in establishing differential and reliable predictors of likely treatment response [@hitchcock_computational_2022]. <!-- MS: how do we know participants are healthy? -->As we found heterogeneity in reactions to a relatively explicit intervention by relatively healthy participants from the general population. is itself heterogeneous as we have shown in this experiment, then the issue of variable treatment responses might the result of the interaction of two sources of variability: the phenotyping of the disorder as well as the phenomenological aspects of the intervention itself. As such, a rigorous exploration of the determinants of inter-individual differences to an intervention in the general patient population is required.
